---
title: "Assignment2-markdown"
author: "Chengwen Luo"
date: "17/03/2018"
output: html_document
keep_md: true
---

## Loading and preprocessing the data
```{r,results='markup'}
library(lubridate)
library(dplyr)
library(ggplot2)
dat<-read.csv("activity.csv")
dat_org<-dat
head(dat)
```

## 2. Histogram of the total number of steps taken each day

``` {r}
tot_step<-as.data.frame(tapply(dat$steps,dat$date,sum))
date<-rownames(tot_step)
rownames(tot_step)=1:dim(tot_step)[1]
tot_step[,2]<-date
names(tot_step)<-c("sum","date")
```

First is a time series bar chat, second one is a histgram
```{r}
g<-ggplot(tot_step,aes(y=sum,x=date))
g+geom_bar(stat="identity") 
hist(tot_step$sum,breaks=61,col="red")
```

### Mean/median/max number of steps taken each day
```{r}
dat$date<-ymd(as.vector(dat$date))
median_step<-tapply(dat$steps,dat$date,median,na.rm=T)
mean_step<-tapply(dat$steps,dat$date,mean,na.rm=T)
max_step<-tapply(dat$steps,dat$date,max)
maxIndex<-which.max(max_step)
max_step[maxIndex] 
head(as.data.frame(median_step))
head(as.data.frame(mean_step))
```

# 2. What is the average daily activity pattern?
Make a time series plot (i.e. ðšðš¢ðš™ðšŽ = "ðš•") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
interval_step<-as.data.frame(tapply(dat$steps,dat$interval,mean,na.rm=T))
interval_step[,2]=rownames(interval_step)
names(interval_step)<-c("steps","interval")
rownames(interval_step)<-1:dim(interval_step)[1]
plot(interval_step$interval,interval_step$steps,type="l",col="red")
```

2 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with ð™½ð™°s)

```{r}
no_NA<-sum(is.na(dat$steps))
no_NA
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
dat_new<-dat
for (i in 1:dim(dat_new)[1]) {
  if (is.na(dat_new[i,1])) {
    na.int_value <- dat_new[i,]$interval
    inter_value_steps<-interval_step[interval_step$interval==na.int_value,]$steps
    dat_new[i,1] = inter_value_steps }
}
```

now calculating the median and mean 

```{r}
new_median<-tapply(dat_new$steps,dat_new$date,median)
new_mean <- tapply(dat_new$steps,dat_new$date,mean)

as.data.frame(new_median)
as.data.frame(new_mean)
```

now calculating the sum of steps per day
```{r}
tot_step2 <- as.data.frame(tapply(dat_new$steps,dat_new$date,sum))
date2 <- rownames(tot_step2)
rownames(tot_step2)=1:dim(tot_step2)[1]
tot_step2[,2]<-date2
names(tot_step2)<-c("sum","date")
```

This is to plot the bar char and histgram
```{r}
g2<-ggplot(tot_step2,aes(y=sum,x=date))
g2+geom_bar(stat="identity") 
hist(tot_step2$sum,col="orange",breaks=61)
```

## Are there differences in activity patterns between weekdays and weekends?
classify workdays and weekends

```{r}
dat_new$weekd<-weekdays(dat_new$date)
for (i in 1:dim(dat_new)[1]) {
   if (dat_new[i,]$weekd %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")) {
     dat_new[i,5] <- "workday"
   } else {dat_new[i,5] <- "weekend"
   }
}
names(dat_new)[5]="weekdayclass"
head(dat_new)
```

now split the data to weekends and weekdays

```{r}
dat_new_split<-split.data.frame(dat_new,dat_new$weekdayclass,drop=F)
interv2_weekend<-as.data.frame(tapply(dat_new_split$weekend$steps,dat_new_split$weekend$interval,mean))
interv2_weekend$interval<-rownames(interv2_weekend)
rownames(interv2_weekend)<-1:dim(interv2_weekend)[1]
interv2_weekend[,3]="weekend"
names(interv2_weekend)<-c("steps","interval","weekclass")


interv2_workday<-as.data.frame(tapply(dat_new_split$workday$steps,dat_new_split$workday$interval,mean))
interv2_workday$interval<-rownames(interv2_workday)
rownames(interv2_workday)<-1:dim(interv2_workday)[1]
interv2_workday[,3]="workday"
names(interv2_workday)<-c("steps","interval","weekclass")
```

plot activity patterns in workdays and weekends

```{r,fig.height=8}


interv_com<-rbind(interv2_workday,interv2_weekend)
interv_com$interval=as.numeric(as.vector(interv_com$interval))
g3<-ggplot(interv_com,aes(interval,steps),colour=factor(weekclass)) + geom_line() + facet_grid(weekclass~.)
g3
```

